The MRChem MRA code for molecular electronic calculations. Performance and scaling properties

### 2022-02-22 main.tex
Updates from Overleaf

\author{Peter Wind, Magnar Bjørve, Anders Brakestad, Luca Frediani, Gabriel Gerez, Stig Rune Jensen, Roberto Di Remigio}

@@

@@

\subsection{Comparison with ORCA REF and LSDalton REF}
\begin{table*}[t]
    \centering
    \begin{tabular}{lllll}
     & E [a.u.] &$\Delta$E [a.u.]& Error [a.u.] & time [s]\\
    LSDalton pc-1 &-3770.20554&-19.12076&0.11&101\\
    LSDalton pc-2 &-3772.56656&-19.27598&-0.04&1688\\
    LSDalton pc-3 &-3772.833116&-19.24894&-0.014&26319\\
    LSDalton$^+$ pc-3 &-3772.82592&-19.26353&-0.029&1924\\
    ORCA pc-1 &-3770.19956&-19.11479&0.12&25\\
    ORCA pc-2 &-3772.56922&-19.27865&-0.044&265\\
    ORCA pc-3 &-3772.83357&-19.24940&-0.014&3933\\
    MRChem mw4&-3772.85028&-19.21937&0.015&117\\
    MRChem mw5&--3772.86560&-19.23469&0.00025& 97$^*$\\
    MRChem mw6&3772.86584&-19.234934&0.00001&251$^*$\\
    MRChem mw7&-3772.86585&-19.234944&0&644$^*$\\
       \end{tabular}    
    \caption{Comparison of results and computation times, for one SCF cycle (Hartree Fock, Valinomycine molecule). LSDalton$^+$ is performed using the accelerators "DENSFIT" and "ADMM". E is the total Hartree fock Energy and $\Delta$E is the atomization energy. The error is defined as the difference with the mw7 results. time is the wall time for one scf cycle. $^*$All computation where done on 4 compute nodes, except for MRChem mw5, mw6 and mw7 which were performed using 16 compute nodes}
        \label{tab:MRC_ORCA_LSD}
\end{table*}

@@

It should be noted that the run times given for ORCA and LSDalton could probably be improved by tuning different input parameters, as we did not try to further optimize all the settings. Still the main picture should not be affected.

@@

Also for DFT methods, it has already been shown (Bischoff 2019) that MRA methods can be competitive wrt computation time.

@@

%to be expanded?
The ability to better control precision is a definitive advantage of the method.

@@

\bibitem[penchoff (2021)]{penchoff2021}https://pubs.acs.org/doi/full/10.1021/bk-2021-1388.ch001  HPC challenges: An Introduction to High Performance Computing and Its Intersection with Advances in Modeling Rare Earth Elements and Actinides

>>>>

\author{Peter Wind, Magnar Bjørve, Anders Brakestad, Gabriel Gerez, Stig Rune Jensen, Roberto Di Remigio and Luca Frediani}

@@

%define precisions mw5, mw6 ...

@@

\subsection{Comparison with ORCA  and LSDalton  }

\begin{table*}[t]
    \centering
    \begin{tabular}{lllll}
     & E [a.u.] &$\Delta$E [a.u.]& Error [a.u.] & time [s]\\
    LSDalton pc-1 &-3770.20554&-19.12076&0.11&101\\
    LSDalton pc-2 &-3772.56656&-19.27598&-0.04&1688\\
    LSDalton pc-3 &-3772.833116&-19.24894&-0.014&26319\\
    LSDalton$^+$ pc-3 &-3772.82592&-19.26353&-0.029&1924\\
    ORCA pc-1 &-3770.19956&-19.11479&0.12&25\\
    ORCA pc-2 &-3772.56922&-19.27865&-0.044&265\\
    ORCA pc-3 &-3772.83357&-19.24940&-0.014&3933\\
    MRChem mw4&-3772.85028&-19.21937&0.015&117\\
    MRChem mw5&--3772.86560&-19.23469&0.00025& 97$^*$\\
    MRChem mw6&3772.86584&-19.234934&0.00001&251$^*$\\
    MRChem mw7&-3772.86585&-19.234944&0&644$^*$\\
       \end{tabular}    
    \caption{Comparison of results and computation times, for one SCF cycle (Hartree Fock, Valinomycine molecule). LSDalton$^+$ is performed using the accelerators "DENSFIT" and "ADMM". E is the total Hartree fock Energy and $\Delta$E is the atomization energy. The error is defined as the difference with the mw7 results. time is the wall time for one scf cycle. $^*$All computation where done on 4 compute nodes, except for MRChem mw5, mw6 and mw7 which were performed using 16 compute nodes}
        \label{tab:MRC_ORCA_LSD}
\end{table*}

@@

It should be noted that the run times given for ORCA\cite{orca} and LSDalton\cite{lsdalton} could probably be improved by tuning different input parameters, as we did not try to further optimize all the settings. Still the main picture should not be affected.

@@

GTO methods have been developed at a time where computer power was a scarce resource. In modern computers the total number of mathematical operations that can be performed is huge. For a complex scientific software, the total number of mathematical operations performed will usually be small compared to the maximum theoretically available on a supercomputer. This is even more true for GPU based computers. Therefore the focus of modern software should shift from minimizing the number of operations, towards finding new ways of partitioning the algorithm in independent parts.

Also for DFT methods, it has already been shown \cite{bischoff2019} MOREREF that MRA methods can be competitive wrt computation time.

@@

The ability to better control precision is a definitive advantage of the method. In methods that use traditional basis sets, the user needs to choose the type of basis, based on what is the property of interest and the precision required. But this is by no mean always a straightforward task and require some degree of expertize. The resulting accuracy is often difficult to assess. What's more the results will suffer from BSSE REF, which complicate both the interpretation and the procedure.
In MRA method, the precision is chosen as a single number that can be chosen on a continuous scale. The resulting precision will be independent of the property of interest. %is this true?

@@

\bibitem[penchoff (2021)]{penchoff2021} https://pubs.acs.org/doi/full/10.1021/bk-2021-1388.ch001  HPC challenges: An Introduction to High Performance Computing and Its Intersection with Advances in Modeling Rare Earth Elements and Actinides

\bibitem[ORCA (2020)]{orca} F. Neese, F. Wennmohs, U. Becker, and C. Riplinger, “The ORCA quantum chemistry program package”, J. Chem. Phys. 152, 224108 (2020). https://doi.org/10.1063/5.0004608

\bibitem[LSDalton (2014)]{lsdalton} K. Aidas, C. Angeli, K. L. Bak, V. Bakken, R. Bast, L. Boman, O. Christiansen, R. Cimiraglia, S. Coriani, P. Dahle, E. K. Dalskov, U. Ekström, T. Enevoldsen, J. J. Eriksen, P. Ettenhuber, B. Fernández, L. Ferrighi, H. Fliegl, L. Frediani, K. Hald, A. Halkier, C. Hättig, H. Heiberg, T. Helgaker, A. C. Hennum, H. Hettema, E. Hjertenæs, S. Høst, I.-M. Høyvik, M. F. Iozzi, B. Jansik, H. J. Aa. Jensen, D. Jonsson, P. Jørgensen, J. Kauczor, S. Kirpekar, T. Kjærgaard, W. Klopper, S. Knecht, R. Kobayashi, H. Koch, J. Kongsted, A. Krapp, K. Kristensen, A. Ligabue, O. B. Lutnæs, J. I. Melo, K. V. Mikkelsen, R. H. Myhre, C. Neiss, C. B. Nielsen, P. Norman, J. Olsen, J. M. H. Olsen, A. Osted, M. J. Packer, F. Pawlowski, T. B. Pedersen, P. F. Provasi, S. Reine, Z. Rinkevicius, T. A. Ruden, K. Ruud, V. Rybkin, P. Salek, C. C. M. Samson, A. Sánchez de Merás, T. Saue, S. P. A. Sauer, B. Schimmelpfennig, K. Sneskov, A. H. Steindal, K. O. Sylvester-Hvid, P. R. Taylor, A. M. Teale, E. I. Tellgren, D. P. Tew, A. J. Thorvaldsen, L. Thøgersen, O. Vahtras, M. A. Watson, D. J. D. Wilson, M. Ziolkowski, and H. Ågren, "The Dalton quantum chemistry program system", WIREs Comput. Mol. Sci. 2014, 4:269–284 (doi:10.1002/wcms.1172)
